Prediction of the Manner of the Exercise
===

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

## Training
### Pre-processing
```{r pre-set}
library(caret)
#setwd('/Users/Nan/Documents/GitRepo/prac_ML/')
```

Load training dataset.

```{r load training,cache=TRUE}
training = read.csv('pml-training.csv',stringsAsFactors=F)
dim(training)

outcome = factor(training$classe)    #save the outcome apart
training = subset(training,select=-classe)
```

There are non-numeric variables that should be numeric, such as "kurtosis_roll_belt". Re-examining the data I found there are records with "#DIV/0!" (probably generated in Excel, orz), which prevented the variable to be read as numeric. Set those to be missing and convert these variables to be numeric.

```{r}
num = sapply(training,is.numeric)
head(num,n = 12)

tmp = as.data.frame(lapply(training[!num],function(x) gsub('#DIV/0!','',x)),stringsAsFactors=F)
dim(tmp)
head(tmp[,1:4])    #the first 3 columns should be still as character
tmp[4:36] = as.data.frame(lapply(tmp[,4:36],as.numeric))    #convert
training[!num] = tmp
```

Remove variables with a missing rate >= 0.95.

```{r}
na_rate = apply(training,2,function(x) sum(is.na(x))/length(x))    #rate of missingness
training = training[,na_rate<0.95]
dim(training)
```

Then remove non-numeric variables and the first column which is the sequence number. Remove near-zero variables.

```{r}
num = sapply(training,is.numeric)
num[1] = FALSE    #also remove the first column
training = training[num]

nzv = nearZeroVar(training,saveMetrics=T)
all(!nzv$nzv)    #none of the variables are near to zero
dim(training)
```

Then standardize the variables.

```{r}
pre_pro1 = preProcess(training,method=c('center','scale','pca'),thresh=0.9)
train_std = predict(pre_pro1,training)
```

Take a subset of training set, since the whole dataset is too big...

```{r}
set.seed(123)

in_train = createDataPartition(y=outcome,p=0.2,list=F)
sub_train = train_std[in_train,]
sub_out = outcome[in_train]
dim(sub_train)
```

### Analyze and evaluate

Splice the data into 10-fold subsets. Train and cross validate.

```{r,cache=TRUE}
fitControl <- trainControl(method="repeatedcv",number=10,repeats=1)    #control of fitting
par_grid <- expand.grid(mtry=c(2,5,10,20))    #parameter grid
model <- train(sub_out~.,data=sub_train,method='rf',trControl=fitControl,tuneGrid=par_grid)
```

The final model is

```{r}
model
model$results
```

with the hyper-parameter mtry=2, and the corresponding **out-of-sample error** is (1-Accuracy), which is `r 1-model$results$Accuracy[1]`.

## Testing
```{r load testing,cache=TRUE}
testing = read.csv('pml-testing.csv')
dim(testing)
```

Pre-process the data as what was done for the training set.

```{r}
testing = testing[names(training)]
dim(testing)

test_std = predict(pre_pro1,testing)
```

Then make predictions.

```{r}
pred = predict(model,test_std)
pred
```